---
title: 'Project 1'
author: "Mora"
date: "15/10/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Henderson and Velleman (1981) comment in a footnote to Table 1: ‘Hocking [original transcriber]'s noncrucial coding of the Mazda's rotary engine as a straight six-cylinder engine and the Porsche's flat engine as a V engine, as well as the inclusion of the diesel Mercedes 240D, have been retained to enable direct comparisons to be made with previous analyses.’

Source: [Source](https://www.jstor.org/stable/2530428?typeAccessWorkflow=login)

```{r cars}
data(mtcars)
summary(mtcars)
```

# Data visualization

## 1. Scatter Plot: mpg vs hp

To study the relationship between Miles per Gallon and the Horsepower I perform an Scatter plot plus adding a smooth line with a shaded area to represent the confidence interval.

```{r pressure, echo=FALSE, fig.align='center'}
library(ggplot2)
ggplot(mtcars, aes(x = mpg, y = hp)) +
  geom_count(color='#FFAB91') +
  geom_smooth(se=TRUE, color='#7E57C2') +
  ggtitle('Miles per gallon vs Gross Horsepower')
```

Notice the plot seems to indicate a negative correlation between hp and mpg. Specially at the range mpg=(13,23) we see as mpg increases, hp decreases with the highest confidence and rate.


## 2. Bar plot: Number of cylinders count


```{r, fig.width=4, fig.height=3,fig.align='center'}
ggplot(data = mtcars) +
  geom_bar(aes(x = factor(cyl), y = ..count..), fill = "#B39DDB", width= 0.5)+
  labs(x='Number of cylinders', y='Count')

```

We can see in the bar plot the most frequent number of cylinders is 8, then 5 and lastly 6


## 3. Box plot: mpg vs gear

To study the relationship between miles per gallon and number of forward gears, I study a Box plot with mpf and groups of gear.
  
```{r, fig.width=5, fig.height=3, echo=FALSE, fig.align='center'}
box_plot_mpg_gear <- ggplot(data = mtcars, aes(x = factor(gear), y = mpg, group = gear, fill = factor(gear))) +
  geom_boxplot() +
  labs(x = 'Gear', y = 'Miles per Gallon',fill = 'Gear Groups') +
  scale_fill_manual(values = c("#CCCCFF", "#6633FF", "#CC99FF"))

box_plot_mpg_gear

```

We notice at 4 and 5 number of forward gears, there is an increase on miles per gallon
So it seems that there is a positive correlation between the different Gear Groups and mpg
However, 4 gears seems to be slightly more increased compared to Gear.

# Statistical Analysis

## 4. Descriptive analysis of mpg

```{r}
mpg_mean <- mean(mtcars$mpg)
cat('mpg mean:', mpg_mean, '\n\n')

mpg_median <- median(mtcars$mpg)
cat('mpg median:', mpg_median, '\n\n')


mpg_sd <- sd(mtcars$mpg)
cat('mpg standard deviation:', mpg_sd)

```

## 5. Correlation analysis between mpg and hp

In the Scatter plot we noticed a negative correlation between mpg and hp. In order to further study this, we look for the correlation coefficient:

```{r}

corr_mpg_hp <- cor(mtcars$hp,mtcars$mpg)
cat('Correlation coefficient between mpg and hp: ',corr_mpg_hp,'\n')

cat('p-value: ',cor.test(mtcars$hp,mtcars$mpg)$p.value)


```


## 6. t-test:  Mean "mpg" of cars with 4 and 6 cylinders

```{r}
with(mtcars, t.test(mpg[cyl==6],mpg[cyl==4]))

```


After performing a t-test to compare the mpg mean between automobiles with 4 and 6 cylinders.
The p-value is < 0.05 and t-value is = -4.7191


## 7. Linear regression to predict mgp based on hp

```{r}
model.fit <- lm(mpg~hp,data=mtcars)
cat('Summary of the model fit:\n')
summary(model.fit)
```

Based on the analysis of the linear regression to predict mpg based on hp, we see with high confidence (p-value <0.05), that the Miles per Gallon can be predicted with the model:

**mpg = 30.09886 -0.06823 · hp + error**
